---
title: "Functionalizing Rasters from NEON"
author: "Naupaka Zimmerman"
date: "June 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

Start by putting up objectives/tasks that students will be working though:

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Visualize density and plot vertical cutoff lines.
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that 6m is generally understory, 10m small trees,and the rest are tall trees. A function could import the desired thresholds. 
1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/ for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

```{r}
# load library and functions
library("raster")

source("scripts/functions.R")

# Setting some parameters
chm_file <- "../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif"
chosen.breaks <- c(6, 30, 40, 50, 60)

chm <- raster(chm_file)

# quick sanity check
plot(chm)

# set 0 values
chm[chm == 0] <- NA
plot(chm)

# check dinsity plot
density(chm)

reclass.height.mat <- create_height_class_matrix(chosen.breaks)


plot_chm_density(rast.in = chm, 
				 title = "Canopy heights at Teakettle", 
				 bins = chosen.breaks)

# reclassify our raster
reclassified.chm <- reclassify(chm,
							   reclass.height.mat)

# plot density figure with breaks to pdf file
make_pdf(plot_chm_density(rast.in = chm, 
				 title = "Canopy heights at Teakettle", 
				 bins = chosen.breaks), 
		 filename = "TEAK_CHM_density_with_breaks.pdf", 
		 width = 6, height = 7)

# Save pdf of reclassified raster
make_pdf(plot_reclassified_raster(reclassified.chm, 
								  site.name = "Teakettle",
								  breaks = chosen.breaks), 
		 filename = "canopy_height_map_reclassified_TEAK.pdf",
		 width = 6, height = 6)

# Save the reclassified raster to a GeoTIFF
writeRaster(reclassified.chm,
			filename = "reclassified_CHM_TEAK_GeoTIFF.tif",
			format = "GTiff",
			options = "COMPRESS=LZW",
			overwrite = TRUE,
			NAflag = -9999)

sink(paste0(format(Sys.time(), "%Y-%m-%d_%H%M%S"),
			"_sessionInfo.txt"))
sessionInfo()
sink()

```
